Notes on: Revision for JMLR Submission on Local Spectral Clustering of Density Upper Level Sets

### 2021-12-23 notes/weekly_notes/10.31.20.NotesForWeek.tex

\documentclass{report}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}
\usepackage{subcaption}
\usepackage{tikz}

@@

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\mathrm{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

@@

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\Xsetx}{\,dx}
\newcommand{\Xsety}{\,dy}
\newcommand{\Xsetr}{\,dr}
\newcommand{\Xsetxpr}{\,dx'}
\newcommand{\Xsetypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LE}{\mathrm{LE}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\OS}{\mathrm{OS}}
\newcommand{\PLS}{\mathrm{PLS}}
\newcommand{\dist}{\mathrm{dist}}

@@

\begin{document}
\title{Work for revision of ``Local Spectral Clustering of Density Upper Level Sets''}
\author{Alden Green}
\date{\today}
\maketitle

\chapter{Normalized cut}
To recall, our upper bound on normalized cut---ignoring the terms that depend on the density---is 
\begin{equation*}
\wt{\Phi}_{\nu,r}(\mc{C}_{\sigma}) \lessapprox \frac{16dr}{9\sigma}.
\end{equation*}
\section{Lower bound on normalized cut}
Our goal is to examine the tightness of our upper bound on normalized cut (Theorem 10 of our JMLR submission), by giving a comparable lower bound. Recall that the (uniform, population level) normalized cut of a set $\mc{C}_{\sigma}$ is given by
\begin{equation*}
\Phi_{\nu,r}(\mc{C}_{\sigma}) := \frac{\int_{\mc{C}_{\sigma}} \nu(B(x,r) \setminus \mc{C}_{\sigma}) \,dx}{\int_{\mc{C}_{\sigma}} \nu(B(x,r)) \,dx}.
\end{equation*}
As informed by the isoperimetric ratio, the worst-case for a functional like normalized cut should be when $\mc{C}_{\sigma}$ by a $d$-dimensional ball. Letting $\mc{C}_{\sigma} = B(0,\sigma)$, we have that
\begin{equation*}
\frac{\int_{\mc{C}_{\sigma}} \nu(B(x,r) \setminus \mc{C}_{\sigma}) \,dx}{\int_{\mc{C}_{\sigma}} \nu(B(x,r)) \,dx} = \frac{1}{\nu_d^2 r^d \sigma^d} \int_{\mc{C}_{\sigma}} \nu(B(x,r) \setminus \mc{C}_{\sigma}) \,dx.
\end{equation*}
If $\|x\| < \sigma - r$, then $B(x,r) \setminus \mc{C}_{\sigma} = \emptyset$. Otherwise $\sigma - r \leq \|x\| \leq \sigma$, and if $r \ll \sigma$ then $B(x,r) \setminus \mc{C}_{\sigma} \approx \mathrm{cap}_{r}(\|x\| - (\sigma - r))$. Converting to polar coordinates, we have:
\begin{align*}
\int_{\mc{C}_{\sigma}} \nu(B(x,r) \setminus \mc{C}_{\sigma}) \,dx & \approx s_d \int_{\sigma - r}^{\sigma} \nu(\mathrm{cap}_{r}(x - (\sigma - r))) t^{d - 1} \,dt \\
& \geq s_d (\sigma - r)^{d - 1} \int_{\sigma - r}^{\sigma} \nu(\mathrm{cap}_{r}(t - (\sigma - r)))  \,dt \\
& = s_d (\sigma - r)^{d - 1} \int_{0}^{r} \nu(\mathrm{cap}_{r}(z)  \,dz.
\end{align*}
As calculated below~\eqref{eqn:lb_conductance_1}, $\int_{0}^{r} \nu(\mathrm{cap}_{r}(z)  \,dz = \frac{1}{d + 1} \nu_d r^{d + 1} \frac{\Gamma(d/2 + 1)}{\Gamma(d/2 + 1/2)\sqrt{\pi}} \approx \frac{1}{d + 1} \nu_d r^{d + 1} \frac{\sqrt{d + 2}}{\sqrt{2\pi}}$ and hence
\begin{equation*}
\Phi_{\nu,r}(\mc{C}_{\sigma}) \approx s_d (\sigma - r)^{d - 1}\frac{1}{d + 1} \nu_d r^{d + 1} \frac{\sqrt{d + 2}}{\sqrt{2\pi}} \frac{1}{\nu_d^2 r^d \sigma^d} \approx \frac{ \sqrt{d + 2}}{(d + 1)\sqrt{2\pi}} \cdot \frac{dr}{\sigma}.
\end{equation*}
By contrast, our upper bound is $\approx dr/\sigma$, revealing a gap of roughly $\frac{ \sqrt{d + 2}}{(d + 1)\sqrt{2\pi}}$.

\chapter{Lower bound for conductance of continuous-space ball walk}
For $x \in \mc{C}_{\sigma}$ and measurable $\Sset \subseteq \mc{C}_{\sigma}$, let $\ell_{\nu,r}(x) = \nu(B(x,r) \cap \mc{C}_{\sigma})$ and ${J}_{\nu,r}(\mc{S}) = \int_{\mc{S}} \ell_{\nu,r}(x) \,dx$. Further define
\begin{equation*}
\widetilde{P}_{\nu,r}(x; \Sset) := \frac{\nu(\Sset \cap B(x,r))}{\nu(\mc{C}_{\sigma} \cap B(x,r))}, \quad \widetilde{\pi}_{\nu,r}(\Sset) = \frac{J_{\nu,r}(\mc{S})}{J_{\nu,r}(\mc{C}_{\sigma})}, \quad  \widetilde{Q}_{\nu,r}(\Sset_1,\Sset_2) := \int_{\Sset_1} \widetilde{P}_{\nu,r}(x;\Sset_2) \,d\widetilde{\pi}_{\nu,r}(x).
\end{equation*}
The uniform continuous normalized cut and conductance are then
\begin{equation*}
\widetilde{\Phi}_{\nu,r}(\Sset) := \frac{\widetilde{Q}_{\nu,r}(\Sset, \Sset^c)}{\min\set{\widetilde{\pi}_{\nu,r}(\Sset),\widetilde{\pi}_{\nu,r}(\Sset^c)}}, \quad \widetilde{\Phi}_{\nu,r} := \min_{\Sset \subseteq \mc{C}_{\sigma}} \widetilde{\Phi}_{\nu,r}(\Sset),
\end{equation*}
where the minimum is taken over all measurable sets $\Sset \subseteq \mc{C}_{\sigma}$.

\begin{lemma}
	\label{lem: uniform_continuous_conductance}
	Let $\mc{C}_{\sigma}$ satisfy Assumption \textcolor{red}{(A3)} (of our main document) for some convex set $\mathcal{K}$ with diameter $\rho$, and measure-preserving mapping $g: \mathcal{K} \to \mc{C}_{\sigma}$ with Lipschitz constant $L$. Let $0 < \ell < 1/2$ be a number that satisfies $\ell_{\nu,r}(x) \geq \nu_d r^d \ell$ for all $x \in \mc{C}_{\sigma}$.  Then the uniform conductance $\widetilde{\Phi}_{\nu,r}$ is lower bounded:
	\begin{equation*}
	\widetilde{\Phi}_{\nu,r} \geq \Bigl(1 - \frac{r}{4\rho}\Bigr) \cdot\frac{\ell^2\sqrt{2\pi}}{9} \cdot \frac{r}{\rho L \sqrt{d + 2}}
	\end{equation*}
\end{lemma}
Lemma~\ref{lem: overlap_balls} gives the lower bound $\ell \geq 1/2 \cdot (1 - r/\sigma \cdot \sqrt{(d + 2)/(2\pi)})$ leading to the following corollary.
\begin{corollary}
	\label{cor: uniform_continuous_conductance}
	Let $\mc{C}_{\sigma}$ satisfy Assumption \textcolor{red}{(A3)} for some convex set $\mathcal{K}$ with diameter $\rho$, and measure-preserving mapping $g: \mathcal{K} \to \mc{C}_{\sigma}$ with Lipschitz constant $L$. Then the uniform conductance $\widetilde{\Phi}_{\nu,r}$ is lower bounded:
	\begin{equation*}
	\widetilde{\Phi}_{\nu,r} \geq \Bigl(1 - \frac{r}{4\rho}\Bigr) \Bigl(1 - \frac{r}{\sigma}\sqrt{\frac{d + 2}{2\pi}}\Bigr)^2 \cdot \frac{\sqrt{2\pi}}{36} \cdot \frac{r}{\rho L \sqrt{d + 2}}
	\end{equation*}
\end{corollary}

\subsection{Proof of Lemma~\ref{lem: uniform_continuous_conductance}}
Lower bounds on the conductance of geometric random walks follow a typical pattern, in which one supplies an isoperimetric inequality, uses this to lower bound the ergodic flow $\wt{Q}$, and thereby derives a lower bound on the conductance. The proof of Lemma~\ref{lem: uniform_continuous_conductance} proceeds along these lines, and we start by sketching the chain of inequalities we will use. For a given partition $\{\mc{S},\mc{R}\}$ of $\mc{C}_{\sigma}$, write
\begin{equation*}
\mc{S}^{\delta} := \biggl\{u \in \mc{S}: \wt{P}(u,\mc{R}) \leq \delta \biggr\}~~\textrm{and}~~\mc{R}^{\delta} := \biggl\{u \in \mc{R}: \wt{P}(u,\mc{S}) \leq \delta \biggr\} 
\end{equation*}
where $\delta$ is a small number we will determine later on. Let $\mc{S}_{\delta} := \mc{C}_{\sigma} \setminus (\mc{S}^{\delta} \cup \mc{R}^{\delta})$. 
\begin{enumerate}
	\item[1.] The ergodic flow $\wt{Q}$ between $\mc{S}$ and $\mc{R}$ satisfies the following lower bound:
	\begin{equation*}
	\wt{Q}_{\nu,r}(\mc{S},\mc{R}) \geq \frac{1}{2} \wt{\pi}_{\nu,r}(\mc{S}_{\delta}) \cdot \delta.
	\end{equation*}
	\item[2.] For any $\mc{S} \subseteq \mc{C}_{\sigma}$, it holds that
	\begin{equation*}
	\frac{\nu(\mc{S})}{\nu(\mc{C}_{\sigma})} \cdot \ell \leq \wt{\pi}_{\nu,r}(\mc{S}) \leq \frac{\nu_d r^d \nu(\mc{S})}{I(\mc{C}_{\sigma})}
	\end{equation*}
	\item[3.] From \textcolor{red}{Abbasi-Yadkori} (see also the seminal works of \textcolor{red}{(Lovasz and Simonovits)} and \textcolor{red}{(Dyer)}), we have the following isoperimetric inequality.
	\begin{lemma}[Isoperimetry of Lipschitz embeddings of convex sets.]
		\label{lem: nonconvex_isoperimetry}
		Let $\mc{C}_{\sigma}$ satisfy Assumption \textcolor{red}{(A3)} for some convex set $\mathcal{K}$ with diameter $\rho$, and measure-preserving mapping $g: \mathcal{K} \to \mc{C}_{\sigma}$ with Lipschitz constant $L$. Then, for any partition $(\Omega_1,\Omega_2,\Omega_3)$ of $\mc{C}_{\sigma}$, 
		\begin{equation*}
		\nu(\Omega_3) \geq 2\frac{\dist(\Omega_1, \Omega_2)}{\rho L} \min(\nu(\Omega_1), \nu(\Omega_2)).
		\end{equation*}
	\end{lemma}
	\item[4.] 
	The following Lemma relates the Euclidean distance $\|x - y\|$ to the the total variation distance
	\begin{equation*}
	\|\wt{P}(x,\cdot) - \wt{P}(y,\cdot)\|_{\mathrm{TV}} = \sup_{\mc{S} \subseteq \mc{C}_{\sigma}} |\wt{P}(x,\mc{S}) - \wt{P}(y,\mc{S}) |
	\end{equation*}
	\begin{lemma}
		\label{lem: one_step_distributions}
		For all $x,y \in \mc{C}_{\sigma}$ and any $\mc{S} \in \mc{C}_{\sigma}$, it holds that
		\begin{equation}
		\label{eqn:one_step_distributions_1}
		\|x - y\| \geq \Bigl(\bigl|\wt{P}(x,\mc{S}) - \wt{P}(y,\mc{S})\bigr|\Bigr)\sqrt{\frac{2\pi}{d + 2}}r\ell
		\end{equation}
		Therefore, for any $\mc{S}_1,\mc{S}_2 \subseteq \mc{C}_{\sigma}$, it holds that
		\begin{equation}
		\label{eqn:one_step_distributions_2}
		\dist(\mc{S}_1,\mc{S}_2) \geq \Bigl(\inf_{\substack{x \in \mc{S}_1 \\ y \in \mc{S}_2}}\|\wt{P}(x,\cdot) - \wt{P}(y,\cdot)\|_{\mathrm{TV}}\Bigr)\sqrt{\frac{2\pi}{d + 2}}r\ell.
		\end{equation}
	\end{lemma}
	Note that for any $x \in \mc{S}^{\delta}, y \in \mc{R}^{\delta}$, we have that
	\begin{equation*}
	\wt{P}(x,\mc{S}) - \wt{P}(y,\mc{S}) = 1 - \bigl(\wt{P}(x,\mc{R}) + \wt{P}(y,\mc{S})\bigr) \geq 1 - 2\delta,
	\end{equation*}
	Thus, applying Lemma~\ref{lem: one_step_distributions} to $\mc{S}^{\delta}$ and $\mc{R}^{\delta}$ gives
	\begin{equation*}
	\dist(\mc{S}^{\delta},\mc{R}^{\delta}) \geq \bigl(1 - 2\delta\bigr) \cdot \sqrt{\frac{2\pi}{d + 2}}r\ell
	\end{equation*}
\end{enumerate}
The claim of Lemma~\ref{lem: uniform_continuous_conductance}---i.e. a lower bound on the continuous normalized cut $\wt{\Phi}_{\nu,r}(\mc{S})$ that holds for any measurable $\mc{S} \subseteq \mc{C}_{\sigma}$---directly follows from these inequalities, as we now show. In particular, the ergodic flow $\wt{Q}(\mc{S},\mc{R})$ can be lower bounded
\begin{equation*}
\wt{Q}_{\nu,r}(\mc{S},\mc{R}) \geq \frac{1}{2}\wt{\pi}_{\nu,r}(\mc{S}_{\delta}) \cdot \delta. \tag{Step 1}
\end{equation*}
Now, suppose $\wt{\pi}_{\nu,r}(\mc{S}^{\delta}) \leq (1 - a) \wt{\pi}_{\nu,r}(\mc{S})$ or $\wt{\pi}_{\nu,r}(\mc{R}^{\delta}) \leq (1 - a) \wt{\pi}_{\nu,r}(\mc{R})$, for some $0 < a < 1$. Then $\wt{\pi}_{\nu,r}(\mc{S}_{\delta}) \geq a \min\{\wt{\pi}_{\nu,r}(\mc{S}),\wt{\pi}_{\nu,r}(\mc{R})\}$, and consequently,
\begin{equation*}
\wt{\Phi}_{\nu,r}(\mc{S}) = \frac{\wt{Q}_{\nu,r}(\mc{S},\mc{S}^c)}{\min\{\wt{\pi}_{\nu,r}(\mc{S}),\wt{\pi}_{\nu,r}(\mc{R})\}} \geq \frac{1}{2}a \delta.
\end{equation*}
Otherwise, we deduce that
\begin{align*}
\wt{Q}_{\nu,r}(\mc{S},\mc{R}) & \geq \frac{\nu(\mc{S}_{\delta})\nu_dr^d}{2I(\mc{C}_{\sigma})} \cdot \ell \delta \tag{Step 2} \\
& \geq \frac{\dist(\mc{S}^{\delta},\mc{R}^{\delta})}{\rho L} \frac{\nu_dr^d\min\{\nu(\mc{S}^{\delta}),\nu(\mc{R}^{\delta})\}}{J(\mc{C}_{\sigma})}  \ell^2 \delta \tag{Step 3}\\
& \geq \frac{r (1 - 2\delta)\sqrt{2\pi}}{\rho L\sqrt{d + 2}} \frac{\nu_dr^d \min\{\nu(\mc{S}^{\delta}),\nu(\mc{R}^{\delta})\}}{J(\mc{C}_{\sigma})}  \ell^2 \delta \tag{Step 4} \\
& \geq  \frac{r (1 - 2\delta)\sqrt{2\pi}}{\rho L\sqrt{d + 2}} \min\bigl\{\wt{\pi}_{\nu,r}(\mc{S}^{\delta}),\wt{\pi}_{\nu,r}(\mc{R}^{\delta})\bigr\} \ell^2 \delta \tag{Step 2} \\
& \geq \frac{r (1 - 2\delta)\sqrt{2\pi}}{\rho L\sqrt{d + 2}} \min\bigl\{\wt{\pi}_{\nu,r}(\mc{S}),\wt{\pi}_{\nu,r}(\mc{R})\bigr\}   \ell^2 \delta (1 - a). \tag{by hypothesis}
\end{align*}
Setting $\delta = 1/3$, we have
\begin{equation*}
\wt{\Phi}_{\nu,r}(\mc{S}) \geq \min\biggl\{\frac{\sqrt{2\pi} r}{9\sqrt{d + 2} \rho L} \ell^2(1 - a), \frac{a\ell}{6}\biggr\}
\end{equation*}
and putting $a = r\ell/(2\rho)$ yields the claim.

Now, we turn to establishing the inequalities used in the above chain.

\paragraph{Step 1: Lower bound the ergodic flow.}
By symmetry, $\wt{Q}_{\nu,r}(\mc{S}_1,\mc{S}_2) = \wt{Q}_{\nu,r}(\mc{S}_1,\mc{S}_2)$, whence a direct computation yields
\begin{equation*}
2\wt{Q}_{\nu,r}(\mc{S}_1,\mc{S}_2) = \wt{Q}_{\nu,r}(\mc{S}_1,\mc{S}_2) + \wt{Q}_{\nu,r}(\mc{S}_2,\mc{S}_1) \geq \delta \wt{\pi}_{\nu,r}(\mc{S}_{\delta}).
\end{equation*}

\paragraph{Step 2: Estimate the stationary distribution.}

Both the upper and lower  bounds on $\wt{\pi}_{\nu,r}(\mc{S})$ are a direct consequence of the following inequalities:
\begin{equation*}
\ell \nu_d r^d \nu(\mc{S})  \leq {J}_{\nu,r}(\mc{S}) \leq \nu_d r^d \nu(\mc{S}).
\end{equation*}
\paragraph{Step 3: Isoperimetric inequality.}
NB: the proof of Lemma \ref{lem: nonconvex_isoperimetry} from first principles is non-trivial, even when the domain $\mc{C}_{\sigma}$ is itself convex, and is a primary technical contribution of the seminal work \textcolor{red}{(Lovasz 1990)}, extended by \textcolor{red}{(Dyer 1991)} among others. However, once the result is shown in the convex setting, it is not hard to show that it applies to Lipschitz transformations of convex sets as well.
\begin{proof}[Proof of Lemma \ref{lem: nonconvex_isoperimetry}]
	For $\Omega_i, i = 1,2,3$, denote the preimage
	\begin{equation*}
	R_i = \set{x \in \mathcal{K}: g(x) \in \Omega_i}
	\end{equation*}
	For any $x \in R_1, y \in R_2$, 
	\begin{equation*}
	\norm{x - y} \geq \frac{1}{L}\norm{g(x) - g(y)} \geq \frac{1}{L} \dist(\Omega_1, \Omega_2). 
	\end{equation*}
	Since $x \in \Omega_1$ and $y \in \Omega_2$ were arbitrary, we have
	\begin{equation*}
	\dist(R_1, R_2) \geq \frac{1}{L} \dist(\Omega_1, \Omega_2).
	\end{equation*}
	By Theorem 2.2 of \textcolor{red}{(Lovasz 1990)},
	\begin{align*}
	\nu(R_3) & \geq 2\frac{\dist(R_1, R_2)}{\rho} \min \{\nu(R_1), \nu(R_2)\} \\
	& \geq \frac{2}{\rho L} \dist(\Omega_1, \Omega_2) \min\{\nu(R_1), \vol(R_2)\}
	\end{align*}
	and by the measure-preserving property of $g$, this implies
	\begin{equation*}
	\nu(\Omega_3) \geq\frac{2}{\rho L} \dist(\Omega_1, \Omega_2) \min\{\nu(\Omega_1), \nu(\Omega_2)\}.
	\end{equation*}
\end{proof}

\paragraph{Step 4: Total variation distance.}

\begin{proof}[Proof of Lemma~\ref{lem: one_step_distributions}]
	Note that~\eqref{eqn:one_step_distributions_2} follows from~\eqref{eqn:one_step_distributions_1} by first taking the supremum over all $\mc{S} \subseteq \mc{C}_{\sigma}$, and then the infimum over all $x \in \mc{S}_1$ and $y \in \mc{S}_2$.

	It remains to show~\eqref{eqn:one_step_distributions_1}. Let $\mc{R} = \mc{C}_{\sigma} \setminus \mc{S}$ and $\mc{I} = B(x,y) \cap B(y,r)$, and suppose without loss of generality that $\ell_{\nu,r}(x) > \ell_{\nu,r}(y)$. We can relate $\wt{P}(x,\mc{S}) - \wt{P}(y,\mc{S})$ to the volume of the intersection $\mc{I}$ as follows:
	\begin{align*}
	\wt{P}(x,\mc{S}) - \wt{P}(y,\mc{S}) & = 1 - \bigl(\wt{P}(x,\mc{R}) + \wt{P}(y,\mc{S})\bigr) \\
	& = 1 - \biggl(\frac{\nu\bigl(B(x,r) \cap \mc{R}\bigr)}{\ell_{\nu,r}(x)} + \frac{\nu\bigl(B(y,r) \cap \mc{S}\bigr)}{\ell_{\nu,r}(y)}\biggr) \\
	& \leq 1 - \frac{1}{\ell_{\nu,r}(x)}\Bigl(\nu\bigl(B(x,r) \cap \mc{R}\bigr) + \nu\bigl(B(y,r) \cap \mc{S}\bigr)\Bigr) \\
	& \leq 1 - \frac{1}{\ell_{\nu,r}(x)}\nu\bigl(\mc{I} \cap \mc{C}_{\sigma}\bigr) \\
	& \leq 1 - \frac{1}{\ell_{\nu,r}(x)}\Bigl(\nu(\mc{I}) - \nu\bigl(B(x,r)\bigr) + \ell_{\nu,r}(x)\Bigr) \\
	& = \frac{1}{\ell_{\nu,r}(x)}\Bigl(\nu_dr^d - \nu(\mc{I})\Bigr).
	\end{align*}
	Now, by hypothesis $\ell_{\nu,r}(x) \geq \ell \nu_dr^d$. On the other hand, Lemma~\ref{lem: volume_of_spherical_cap} gives a lower bound on the volume of $\mc{I}$, and from this we obtain
	\begin{equation*}
	\wt{P}(x,\mc{S}) - \wt{P}(y,\mc{S}) \leq \frac{\|x - y\|}{\ell r} \sqrt{\frac{d + 2}{2\pi}};
	\end{equation*}
	solving for $\|x - y\|$ yields~\eqref{eqn:one_step_distributions_1}.
\end{proof}

\section{Volume Estimates}

>>>>

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{fullpage}
\usepackage{mathtools}
\usepackage{subcaption}
\usepackage{tikz}

@@

\newcommand{\emC}{C_n}
\newcommand{\emCpr}{C'_n}
\newcommand{\emCthick}{C^{\sigma}_n}
\newcommand{\emCprthick}{C'^{\sigma}_n}
\newcommand{\emS}{S^{\sigma}_n}
\newcommand{\estC}{\widehat{C}_n}
\newcommand{\hC}{\hat{C^{\sigma}_n}}
\newcommand{\vol}{\text{vol}}
\newcommand{\spansp}{\mathrm{span}~}
\newcommand{\1}{\mathbf{1}}

@@

%%% Misc
\newcommand{\grid}{\mathrm{grid}}
\newcommand{\critr}{R_n}
\newcommand{\Xsetx}{\,dx}
\newcommand{\Xsety}{\,dy}
\newcommand{\Xsetr}{\,dr}
\newcommand{\Xsetxpr}{\,dx'}
\newcommand{\Xsetypr}{\,dy'}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\spec}{\mathrm{spec}}
\newcommand{\LE}{\mathrm{LE}}
\newcommand{\LS}{\mathrm{LS}}
\newcommand{\OS}{\mathrm{OS}}
\newcommand{\PLS}{\mathrm{PLS}}

@@

\begin{document}
\title{Notes on: Revision for JMLR Submission on Local Spectral Clustering of Density Upper Level Sets}
\author{Alden Green}
\date{\today}
\maketitle

Recall: our Theorem 5 (JMLR numbering) says that with high probability, the following statement holds: when the PPR algorithm is properly initialized, the resulting estimator $\wh{C}$ of an empirical density cluster $\mc{C}_{\sigma}[X]$ satisfies
\begin{equation}
\label{eqn:volume_ssd_ppr}
\frac{\Delta(\wh{C},C_{\sigma}[X])}{\vol_{n,r}(\mc{C}_{\sigma}[X])} \leq c \cdot \kappa(\mc{C}).
\end{equation}
Both reviewers and the editor treated Theorem~5 as our main result, and viewed it as too weak, in related but slightly different ways. Specifically, they worried that:
\begin{itemize}
	\item \textbf{Reviewer 1.} For an example provided in our experiments section, the right hand side is (many times) larger than $1$. Since the trivial estimator $\wh{C} = \emptyset$ will always have
	\begin{equation*}
	\frac{\Delta(\emptyset,C_{\sigma}[X])}{\vol_{n,r}(\mc{C}_{\sigma}[X])}  = 1,
	\end{equation*}
	our bound is vacuous.
	\item \textbf{Reviewer 2.} The right hand side of~\eqref{eqn:volume_ssd_ppr} does not go to $0$ as $n \to \infty$. 
	\item \textbf{Editor.} ``...either the three of us did not get the 
	point of Theorem 5, in which case the presentation
	is obviously not satisfying, or Theorem 5 is not
	as strong as one would like to have it."
\end{itemize}

Both Reviewer 1 and Reviewer 2 make accurate statements, motivating the following pair of research questions:
\begin{enumerate}
	\item \textbf{Addressing the concerns of Reviewer 1.} What are some distributions $P$ and resulting density clusters $\mc{C}$ for which our bound is not vacuous, i.e. for which
	\begin{equation*}
	c \cdot \kappa(\mc{C}) < 1?
	\end{equation*}
	\item \textbf{Addressing the concerns of Reviewer 2.} Should we expect the right hand side of~\eqref{eqn:volume_ssd_ppr} to go to $0$ as $n \to \infty$?
\end{enumerate}
We address these concerns in the following two sections.

\section{Concerns of Reviewer 1}

\subsection{Uniform rectangles + $\epsilon$-background noise model.}
Let us recall the model used in the construction of our ``hard case''. Letting $\theta = (\epsilon,w,h)$, the distribution $P = P_{\theta}$ is a mixtures of uniforms over the unit square $[-1,1]^2$:
\begin{equation*}
P_{\theta} := \frac{1 - \epsilon}{2} P_{(1)} + \frac{1 - \epsilon}{2} P_{(2)} + \epsilon Q;
\end{equation*}
here $Q$ is the uniform distribution over $[-1,1]^2$, and the mixture component $P_{(i)}$ is the uniform distribution over the rectangle $\mc{C}_{(i)}$, 
\begin{equation*}
\mc{C}_{(i)} := c_{(i)} + \frac{1}{2}\Bigl([-w,w] \times [-h,h]\Bigr)
\end{equation*}
with $c_{(1)} = (-.5,0)$ and $c_{(2)} = (.5,0)$, and. Letting $f_{\theta}$ be the density function of $P_{\theta}$---which has the simple closed form expression
\begin{equation*}
f_{\theta}(x) =
\begin{cases}
\frac{(1 - \epsilon)}{2wh} + \frac{1}{4}\epsilon,& ~~ x \in \mc{C}_{(i)} \\
\frac{1}{4}\epsilon,& ~~ x \in [-1,1]^2 \setminus \bigl(\mc{C}_{(1)} \cup \mc{C}_{(2)}\bigr)
\end{cases}
\end{equation*}
---it is clear that $\mc{C}_{(1)}$ and $\mc{C}_{(2)}$ have higher density than the rest of $[-1,1]^2$, and in particular that for any $\epsilon/4 < \lambda \leq (1 - \epsilon)/(2wh) + \epsilon/4$ we have $\mathbb{C}_f(\lambda) = \{\mc{C}_{(1)}, \mc{C}_{(2)}\}$.

\paragraph{Our bound.}
To be more explicit, our bound upper bound on the volume of the symmetric set difference is given by
\begin{equation}
\label{eqn:volume_ssd_ppr_explicit}
\frac{\Delta(\wh{C},\mc{C}_{\sigma}[X])}{\vol_{n,r}(\mc{C}_{\sigma}[X])} \leq 2180 \cdot \Phi_u(\mc{C}) \cdot \tau_{u}(\mc{C}),
\end{equation}
i.e. the constant in~\eqref{eqn:volume_ssd_ppr} is $c = 2180$, and the condition number $\kappa(\mc{C})$ is equal to the product of (upper bounds on) the normalized cut $\Phi_u(\mc{C})$ and mixing time $\tau_u(\mc{C})$. Technically speaking the density clusters $\mc{C}_{(i)}$ do not satisfy our assumptions on density clusters $\mc{C}_{\sigma}$, but modulo some additional bookkeeping we can compute the following upper bounds:
\begin{equation*}
\Phi_u(\mc{C}) = 
\end{equation*}

\paragraph{}

\begin{itemize}
	\item There exist choices of $\theta$ for which:
	\begin{enumerate}[(i)]
		\item  Our bound on 
		\begin{equation*}
		\frac{\Delta(\wh{C},\mc{C}[X])}{\vol_{n,r}(\mc{C}[X])}
		\end{equation*}
		will be less than $1$. 
		\item 
		\begin{equation*}
		\frac{\Delta(\wh{C},\mc{C}[X])}{\vol_{n,r}(\mc{C}[X])} < 1,
		\end{equation*}
		but any bound based on the results of \textcolor{red}{(Allen-Zhu)} will be $ > 1$.
	\end{enumerate}
	The point (i) is an obvious response to Reviewer 1. The concern is that the values of $\theta$ for which (i) holds are implausible, for instance if $\epsilon$ must be absurdly small. Point (ii) addresses this concern, giving the values of $\theta$ one should expect to derive for if one must use the results of \textcolor{red}{(Allen-Zhu)}. 
	\item \textbf{Rates of convergence in $\epsilon$.} As $\epsilon \to 0$, it is obvious that PPR will come to consistently recover $\mc{C}$. Our upper bound correctly reflects this fact. The following question is perhaps more interesting: does our upper bound correctly capture the rate of convergence?
\end{itemize}

\subsection{Other models?}

\textcolor{red}{(Discuss with Siva.)}

\section{Concerns of Reviewer 2}
We first point out that to the best of our knowledge, all other works on spectral clustering on random geometric graphs fall into three categories. 
\begin{enumerate}[(1)]
	\item Define clusters with respect to geometric conditions, obtain upper bound on classification error that does not go to $0$ as $n \to \infty$.
	\item Obtain consistency, but for clusters defined with respect to a continuum PDE or eigenfunction problem.
	\item Obtain consistency for geometrically defined clusters, by changing the algorithm.
\end{enumerate} 
Our work falls into category (1), and so trades a weaker result---non-vanishing upper bound on classification error---for a more natural definition of cluster.

Of course, simply because other works in category (1) share the non-vanishing upper bound property does not mean that such a property is fundamental to the problem. We investigate the question empirically.

\subsection{Uniform rectangles + $\epsilon$-background noise model}

Let $X_1,\ldots,X_n$ be drawn from $P_{\theta}$, with $w = .5, h = 1$ and $\epsilon \approx .72$. In Figure~\ref{fig:fig1}, we plot the error---volume of the symmetric set difference, normalized by the volume of $\mc{C}_{(1)}$--- of PPR clustering against the sample size $n$. We see that PPR--even when optimally seeded, tuned, and with the optimal sweep cut taken---fails to consistently recover $\mc{C}_{(1)}$. 
\begin{figure}
	\centering
	\begin{subfigure}{.5\textwidth}
	\centering
	\includegraphics[width=.8\linewidth]{figures/11_05_2020/error.pdf}
	\end{subfigure}%
	\begin{subfigure}{.5\textwidth}
		\centering
		\includegraphics[width=.8\linewidth]{figures/11_05_2020/error_upper_bound.pdf}
	\end{subfigure}
	\caption{Error---$\Delta(\wh{C},\mc{C}[X])$ divided by $\vol_{n,r}(\mc{C}_{(1)}[X])$---of PPR clusters $\wh{C}$, averaged over 5 iterations. Different curves correspond to different values of radius $r$. For each value of radius $r$, a range of teleportation parameters $\alpha$ and sweep cuts was considered; we have plotted the smallest error among these ranges.}
	\label{fig:fig1}
\end{figure}

In Figures 2 and 3, we see what the set is that PPR actually recovers. The lesson is unsurprising: PPR trades off geometry for density, preferring an estimate $\wh{C}$ which is somewhat more geometrically well-behaved (i.e lower aspect ratio) even at the expense of including some low-density areas.
\begin{figure}
	\centering
	\includegraphics[width = .5\textwidth]{figures/11_05_2020/ppr_cluster.pdf}
	\caption{PPR cluster estimate, for different values of $r$.}
	\label{fig:fig2}
\end{figure}
